\section{Implications of Nonclumpiness} \label{corollaries}
It is a basic property of the parallel chip-firing game that every vertex fires
the same number of times each period~\cite{jiang}. This means, roughly
speaking, that every periodic game is either ``mostly waiting'' with bursts of
firing or ``mostly firing'' with bursts of waiting. (In fact, there is a
bijection between them. Each periodic game has a complement that inverts firing
and waiting~\cite{jiang}.) This is because if a vertex waits twice in a row,
then because it therefore never fires twice in a row, it fires less than half
the time over the course of a period. Similarly, a vertex that fires twice in a
row fires more than half the time. We cannot have a vertex that waits twice
in a row and a vertex that fires twice in a row in the same periodic game
because each vertex fires the same number of times each period.

\begin{corollary}
Once a parallel chip-firing game reaches a periodic position, either no vertex
fires twice in a row or no vertex waits twice in a row.
\end{corollary}

That is, in periodic games, a firing sequence is possible if and only if it is
nonclumpy.

Let the \emph{interior} of a set of vertices $W$ be $I(W) = \set{v \in W}{N(v)
  \subseteq W}$. Because a waiting (or firing) vertex with only waiting (or
firing) neighbors will wait (or fire) the following turn as well, the above
observation proves the following conjecture of Fey and
Levine~\cite{privateComms}.

\begin{corollary}\label{feyLevine}
Let $\sigma$ be a periodic game on $G$,
\begin{align*}
  A &= \set{v \in \V}{\firing{v}{a} = 0} \\
  B &= \set{v \in \V}{\firing{v}{b} = 1},
\end{align*}
where $a,b \in \nats$. Then either $I(A)$ or $I(B)$ is empty.
\end{corollary}

Interestingly, Corollary~\ref{feyLevine} also implies Theorem~\ref{nct}. If
clumpy PFPs were possible, then a leaf attached to a motor with a clumpy PFP
would be in $I(A)$ and $I(B)$ for appropriately chosen $a,b \in \nats$.

In one of the first papers on the parallel chip-firing game, Bitar and Goles
characterized parallel chip-firing games on trees in~\cite{bitarGoles}.
Corollary~\ref{freeLunch} and Theorem~\ref{nct} allow us to characterize the
behavior on tree-like subgraphs---subgraphs such that, if an edge to a root
vertex is cut, become a tree separated from the rest of the graph---by making
the root vertex a motor.

\begin{corollary}
Let $\s$ be a periodic game on $G$ with $\period \geq 3$ in which no vertex
fires twice in a row, $H$ be a tree-like subgraph of $G$ and $m \in \V[H]$ be
the root of $H$. Then for all $v \in \V[H]$,
\[
  \chips{v}{t} = \begin{cases}
    \deg{v} & \textnormal{if } \firing{m}{t-D} = 1 \\
    0 & \textnormal{if } \firing{m}{t-D-1} = 1 \\
    \deg{v} - 1 & \textnormal{otherwise},
  \end{cases}
\]
where $D$ is the distance from $m$ to $v$. An analogous result holds if no
vertex waits twice in a row.
\end{corollary}

In some sense, tree-like subgraphs are passive in that their vertices fire only
in response to their root-side neighbor firing. In a periodic game, we can
completely remove tree-like subgraphs without affecting the PFPs of the other
vertices.

\begin{corollary}
Let $\s$ be a periodic game on $G$ and $l \in \V$ be a leaf with $\N{l} =
\{m\}$. Suppose that no vertex fires twice in a row. Let $G'$ be the subgraph
of $G$ induced by $\V \setminus \{l\}$. Define $\s'$ on $G'$ such that
$\chips[\s']{v}{0} = \chips{v}{0}$ for all $v \in \V \setminus \{m\}$. Then
\[
  \chips[\s']{m}{0} = \begin{cases}
    \chips{m}{0} &\textnormal{if }\showhere{\firing{l}{0}} = 1 \\
    \chips{m}{0} - 1 &\textnormal{otherwise}.
  \end{cases}
\]
An analogous result holds if no vertex waits twice in a row.
\end{corollary}

Compared to $\s'$, $m$ has to have an extra chip to fire in $\s$. However,
unless $m$ fired the previous turn---which, because $l$ is a leaf, is
equivalent to saying $l$ is firing this turn---$m$ will have received the extra
chip back from $l$, so removing both $l$ and the chip has no effect on $m$ as
long as $m$ does not fire while $l$ has a chip. This corollary concerns a leaf,
though the result generalizes to all tree-like subgraphs by repeated
application, providing an alternate proof of Corollary~\ref{freeLunch}.
